---
title: "pre-class 03"
author: "Nana Boame"
date: "September 22, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Pre-Class 03

We will work with the dataset called gapminder, this is a cleaned up version from Gapminder Data. Gapminder contains a lot of great data on all of the nations of the world. We first need to install the gapminder package in R.

```{r}
install.packages("gapminder")
library(dplyr)
library(gapminder)
gapminder
ls(gapminder)
```

Pre-Class Problems
Use dplyr functions to address the following questions:

How many unique countries are represented per continent?
```{r}
gapminder %>% 
  group_by(continent) %>% #group countries by continent
  summarise(Unique_Countries = n_distinct(country)) #count the number of countries by continent
```
    Africa, the Americas, Asia, Europe and Oceania have 52, 25, 33, 30 and 2 countries, respectively.


Which European nation had the lowest GDP per capita in 1997?
```{r}
gapminder %>%
  filter(year == "1997" & continent == "Europe") %>%  #filter information for European nations in 1997
  arrange(by = gdpPercap) #arrange GDP from lowest to highest
```
  Albania was the European nation with the lowest GDP in 1997. 
  
According to the data available, what was the average life expectancy across each continent in the 1980s?
```{r}
gapminder %>%
  filter(between(year, 1980,1989)) %>%  #filter for data from 1980 to 1989
  group_by(continent) %>% #then group by continent
  summarise(mean_LifeExp = mean(lifeExp)) #then calculate the mean life expectancy
```
  The average life expectancy (in years) across each continent in the 1980s is as follows:
    Africa    52.5
    Americas  67.2
    Asia      63.7
    Europe    73.2
    Oceania   74.8

What 5 countries have the highest total GDP over all years combined?
```{r}
gapminder %>%
  group_by(country) %>% #group by country
  summarise(Total_GDP = sum(gdpPercap)) %>% #then calculate total GDP
  arrange(by = desc(Total_GDP)) #then order by descending total GDP

```
    The 5 countries with the highest total GDP over all years combined are 1) Kuwait, 2) Switzerland, 3) Norway, 4) United States and 5) Canada. 

What countries and years had life expectancies of at least 80 years? N.b. only output the columns of interest: country, life expectancy and year (in that order).
```{r}
gapminder %>%
  filter(lifeExp >= 80) %>% #choose only those rows where life expectancy is equal to or greater than 80
  select(country, lifeExp, year) #select the desired columns
```
  The following countries had life expectancies of at least 80 years in the corresponding years:
    Australia               2002, 2007
    Canada                        2007
    France                        2007
    Hong Kong, China  1997, 2002, 2007
    Iceland                 2002, 2007
    Israel                        2007
    Italy                   2002, 2007
    Japan             1997, 2002, 2007
    New Zealand                   2007
    Norway                        2007
    Spain                         2007
    Sweden                        2007
    Switzerland             2002, 2007

What 10 countries have the strongest correlation (in either direction) between life expectancy and per capita GDP?
```{r}

```


Which combinations of continent (besides Asia) and year have the highest average population across all countries? N.b. your output should include all results sorted by highest average population. With what you already know, this one may stump you. See this Q&A for how to ungroup before arrangeing. This also behaves differently in more recent versions of dplyr.
```{r}

```


Which three countries have had the most consistent population estimates (i.e. lowest standard deviation) across the years of available data?
```{r}

```


Subset gm to only include observations from 1992 and store the results as gm1992. What kind of object is this?
Which observations indicate that the population of a country has decreased from the previous year and the life expectancy has increased from the previous year? See the vignette on window functions.
```{r}

```

