# pre-class


Make sure you commit this often with meaningful messages. 



```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


1. Read the source code for each of the following three functions, puzzle out what they do, and then brainstorm better names.

```{r}
# This function checks if each element in vector 'nchar' begins with the string 'prefix'
check_prefix <- function(string, prefix) {
  substr(string, 1, nchar(prefix)) == prefix
}

# This function removes the last element of the vector
remove_last <- function(x) {
  if (length(x) <= 1) return(NULL)
  x[-length(x)]
}

# This function repeats 'y' for each element of 'x'
repeat_y <- function(x, y) {
  rep(y, length.out = length(x))
}
```

2. Compare and contrast rnorm() and MASS::mvrnorm(). How could you make them more consistent?
```{r}
?rnorm
?MASS::mvrnorm

# rnorm takes samples from the univariate normal distribution. The arguments for this function are n, mean and sd.

# MASS::mvrnorm samples from the multivariate normal distribution. The arguments for this function are n, mu and Sigma.

# These functions could be made consistent by using the same names for the arguments although I'm not completely sure that would be appropriate as sd and sigma are not interchangeable.
```



3. Use `lapply()` and an anonymous function to find the coefficient of variation (the standard deviation divided by the mean) for all columns in the mtcars dataset. 

4. Use vapply() to:
    a. Compute the standard deviation of every column in a numeric data frame.
    b. Compute the standard deviation of every numeric column in a mixed data frame. (Hint: you'll need to use vapply() twice.)